\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-05-11 slides/slides.tex
Slides for Agda Implementors' Meeting XXV
* [ Makefile ] type-checked impl-11.agda.

* [ slides ] init.

* [ slides ] tstp, tptp explanation.

* [ slides ] playing with fonts.

* [ slides ] images PDF files added.

* [ slides ] put some order.

* [ slides ] updated files.

* [ slides ] tptp, and tstp slides added.

* [ slides ] diagrams added.

* [ slides ] clickeable items in the diagram.

* [ agda-prop ] updated.

* [ agda-prop ] fixed import of Weakening module.

* [ slides ] fixed typos.

* [ slides ] fixed mistakes.

* [ slides ] minor changes.

* [ slides ] minor changes.

* [ slides ] M

@@

  \begin{itemize}
  \item Is a language\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}.}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture
      \item[\texttt{formula}] formula in \texttt{TPTP} format
    \end{itemize}
  \end{itemize}

\gobackreconstruction
\end{frame}

@@

\texttt{Metis} is an automatic theorem prover for First-Order Logic with Equality
\citep{hurd2003first}
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in
       \texttt{Standard ML}
    \item Each refutation step is one of
      \hyperlink{metis-inference-rules}{\textit{six rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs \textit{detailed} proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\gobackreconstruction
\end{frame}

@@

where \texttt{\color{plum} source} typically is an inference record
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents).
\end{tptp}
}
\end{center}
\gobackreconstruction
\end{frame}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof problem.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=tstp-dag]{DAG Example}
\vfill
By refutation, we proved $p ⊢ p$:
\begin{center}
\begin{prooftree}
\AxiomC{$p$}
\RightLabel{\texttt{\scriptsize assume}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize strip}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize negate}}
\UnaryInfC{$¬ p$}
\AxiomC{$p$}
\RightLabel{\texttt{\scriptsize canonicalize}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize simplify}}
\BinaryInfC{$⊥$}
\RightLabel{\texttt{\scriptsize canonicalize}}
\UnaryInfC{$⊥$}

\end{prooftree}
% \includegraphics[height=0.8\textheight]{figures/derivation.png}
\end{center}
\vfill
\gobackreconstruction
\end{frame}

@@

\end{itemize}
\gobackreconstruction
\end{frame}

>>>>

\begin{frame}{Obstacles to Reconstruction with Automatic Provers \citep{Paulson2007}}
\begin{itemize}
\item \textit{Ambiguities}: their output typically omits crucial information,
such as which term is affected by rewriting.
\item \textit{Lack of standards}: automatic provers generate different output formats
and employ a variety of inference systems
\item \textit{Complexity}: a single automatic prover may use numerous inference rules with
complicated behaviors
\item \textit{Problem transformations}: ATPs re-order literals and make other changes to the clauses
they are given
\end{itemize}

\end{frame}

@@

  \begin{itemize}
  \item Is a language\footnote{\url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}.}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture
      \item[\texttt{formula}] formula in \texttt{TPTP} format
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\texttt{Metis} is an automatic theorem prover for First-Order Logic with Equality
\citep{hurd2003first}
\begin{block}{Why Metis?}
  \begin{itemize}
    \item \href{https://github.com/gilith/metis}{Open source} implemented in
       \texttt{Standard ML}
    \item Each refutation step is one of
      \hyperlink{metis-inference-rules}{\textit{six rules}}
    \item Reads problem in \hyperlink{tptp-syntax}{\texttt{TPTP} format}
    \item Outputs \textit{detailed} proofs in \hyperlink{tstp-syntax}{\texttt{TSTP} format}
  \end{itemize}
\end{block}
\end{frame}

@@

where \texttt{\color{plum} source} typically is an inference record
\begin{center}
{\footnotesize
\begin{tptp}
inference(rule, useful info, parents).
\end{tptp}
}
\end{center}
\end{frame}

@@

\begin{itemize}
  \item Proof found by \texttt{Metis} for the problem $p ⊢ p$
{\small
\begin{tptp}
$ metis --show proof problem.tptp
fof(a, axiom, p).
fof(goal, conjecture, p).
fof(subgoal_0, plain, p),
  inference(strip, [], [goal])).
fof(negate_0_0, plain, ~ p,
  inference(negate, [], [subgoal_0])).
fof(normalize_0_0, plain, ~ p,
  inference(canonicalize, [], [negate_0_0])).
fof(normalize_0_1, plain, p,
  inference(canonicalize, [], [a])).
fof(normalize_0_2, plain, $false,
  inference(simplify, [],
    [normalize_0_0, normalize_0_1])).
cnf(refute_0_0, plain, $false,
    inference(canonicalize, [], [normalize_0_2])).
\end{tptp}
}
\end{itemize}
\end{frame}

\begin{frame}[fragile, label=tstp-dag]{DAG Example}
\vfill
By refutation, we proved $p ⊢ p$:

\begin{columns}
\begin{column}{0.5\textwidth}

\begin{prooftree}
\AxiomC{}
\RightLabel{\texttt{\scriptsize assume}}
\UnaryInfC{$¬ p$}
\RightLabel{\texttt{\scriptsize strip}}
\UnaryInfC{$¬ p$}
\AxiomC{}
\RightLabel{\texttt{\scriptsize axiom}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize canonicalize}}
\UnaryInfC{$p$}
\RightLabel{\texttt{\scriptsize simplify}}
\BinaryInfC{$⊥$}
\RightLabel{\texttt{\scriptsize canonicalize}}
\UnaryInfC{$⊥$}
\end{prooftree}
\end{column}
\begin{column}{0.5\textwidth}  %%<--- here
    \begin{center}
    \includegraphics[height=0.8\textwidth]{figures/derivation.png}
     \end{center}
\end{column}
\end{columns}
\vfill
\gobackreconstruction
\end{frame}

@@

\end{itemize}
\end{frame}
